---
title: "Water_soluble_and_acidity"
author: "Alex"
date: "2024-06-13"
output: html_document
---


```{r library}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
#Additionally has funcitons like "clean_names"
library(janitor)
#Additionally has functions for ggplot, such as "facet_wrap_paginate"
library(ggforce)
library(factoextra)
library(ggsignif) #Helps add significance relationships
library(GGally)
#To reshape the HCL_Water table
library(reshape2)

```

###Data cleaning of the original excel files
```{r data cleanin}

#Import the files
#2022_water_extract is the one with the water soluble concentrations for all analyzed sites
#Acidity_Final is the one with the acidity measurements for all analyzed sites
#Geochemistry2B contains the HCL and Aqua Regia (AR) concentrations
#Site_summary contains the north-south area division, for merging with the water-extracts

###Precleaning of Water sol
Water_sol_all <- X2022_water_extract[-c(1,184:185), -c(1,3,4)]
colnames(Water_sol_all) <- c("Site", "Triplicate", "Layer", "Treatment", "Al", "As", "B", "Ba", "Be", "Ca", "Cd", "Co", "Cr", "Cu", "Fe", "Hg", "K","Mg", "Mn","Mo", "Na", "Ni", "P", "Pb", "S", "Sb", "Se", "Si", "Sn", "Sr", "Tl", "U", "V", "Zn")
Water_sol_all

Water_sol <- Water_sol_all %>%
  select(1:4, sort(current_vars()))
Water_sol$Layer <- sapply(Water_sol$Layer, as.character)
Water_sol$Layer <- paste(Water_sol$Layer, Water_sol$Treatment)
Water_sol
Water_sol_selected <- Water_sol[,c("Site", "Triplicate", "Layer", "Al", "As", "Cd", "Co", "Cr", "Cu", "Fe", "Mn", "Ni", "Pb", "S", "Zn")]
Water_sol_selected[Water_sol_selected$Triplicate == "A+B+C",2] = "A"
Water_sol_selected[Water_sol_selected$Triplicate == "ABC",2] = "A"
Water_sol_selected[Water_sol_selected$Triplicate == "A+B",2] = "A"
#manually remove S18 RZ and TZ, since it's not hypersulfidic material
Water_sol_selected <- Water_sol_selected[-c(113, 91 ,168:169),]
Water_sol_selected[Water_sol_selected$Site == "S18",]


###Cleaning Acidity
Acidity <- Acidity_Final[-c(163:168),-c(1:3, 7, 11, 12)] #nr 9 is the acidity for pH 5.5 mmol, in case you want it back
Acidity
colnames(Acidity)[1:6] = c("Site", "Layer", "Triplicate", "Treatment", "Acidity1", "Acidity")
Acidity$Layer <- paste(Acidity$Layer, Acidity$Treatment)
Acidity$Layer <- gsub("1 Non-inc", "1 No", Acidity$Layer)
Acidity$Layer <- gsub("2 Non-inc", "2 No", Acidity$Layer)
Acidity$Layer <- gsub("2 Incubated", "2 Yes", Acidity$Layer)
Acidity$Layer <- gsub("3 Incubated", "3 Yes", Acidity$Layer)
Acidity <- Acidity[! duplicated(Acidity[,1:4]),]
Acidity
#remove S18
Acidity <- Acidity[-c(128:131),]
Acidity[Acidity$Triplicate == "A+B+C",3] = "A"
Acidity[Acidity$Triplicate == "ABC",3] = "A"
Acidity[Acidity$Triplicate == "A+B",3] = "A"
Acidity[Acidity$Triplicate == "A+C",3] = "A"
Acidity

#Basic statistics, median, minimum and maximum values extracted
Acidity
Acidity_summary <- Acidity[,c(2,5,6)] %>%
  group_by(Layer) %>%
  summarise_each(funs(median(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE)))
write.csv(Acidity_summary, "Acidity_summary_med.csv")


###Combine Acidity and Water_sol_selected
Water_sol_selected <- full_join(Water_sol_selected, Acidity, by = c("Site", "Layer", "Triplicate"))
Water_sol_selected$treatment2 <- Water_sol_selected$Layer
Water_sol_selected$treatment2 <- gsub("1 No", "Non-incubated", Water_sol_selected$treatment2)
Water_sol_selected$treatment2 <- gsub("2 No", "Non-incubated", Water_sol_selected$treatment2)
Water_sol_selected$treatment2 <- gsub("2 Yes", "Incubated", Water_sol_selected$treatment2)
Water_sol_selected$treatment2 <- gsub("3 Yes", "Incubated", Water_sol_selected$treatment2)
Water_sol_selected <-  Water_sol_selected[,-16]
colnames(Water_sol_selected)[18] <- "Treatment"
Water_sol_selected <- Water_sol_selected[!duplicated(Water_sol_selected[,1:4]),]


#Water and acidity in long format
WALONG <- Water_sol_selected %>%
  gather(key = "Element", value = "mgKg", -"Triplicate", -"Treatment", -"Acidity1", -"Site", -"Layer")

WALONG$Layer <- gsub("1 No", "SUL", WALONG$Layer)
WALONG$Layer <- gsub( "2 No", "TZ", WALONG$Layer)
WALONG$Layer <- gsub( "2 Yes", "TZ (inc)",WALONG$Layer)
WALONG$Layer <- gsub("3 Yes", "HYP",WALONG$Layer)

#Water in long format, with acidity as a separate column
WALONG2 <- Water_sol_selected %>%
  gather(key = "Element", value = "mgKg", -"Triplicate", -"Site", -"Treatment", -"Layer", -"Acidity1", -"Acidity")
WALONG2


##### HCL 
HCL_upd <- Geochemistry2B_HCL
HCL_upd <- HCL_upd[-c(1,2,201:269), -c(1,2,4,7)]
colnames(HCL_upd)[1:3] <- c("Triplicate", "Site", "Layer") 
HCL_upd <- HCL_upd %>%
  separate(Triplicate, c("deleteme", "Triplicate"), -1)
HCL_upd <- HCL_upd[,c("Triplicate", "Site", "Layer", "Al", "As", "Cd", "Co", "Cr", "Cu", "Fe", "Mn", "Ni", "Pb", "S", "Zn")]
HCL_upd[,4:15] <- lapply(HCL_upd[,4:15], as.numeric)
HCL_upd

#### Aqua Regia

AR <- Geochemistry2B_Aqua[-c(1,2, 201:223),c(3,5,9, 35)]
colnames(AR) <-  c("Temp", "Site", "Layer", "S")
AR <- separate(AR, "Temp", into = c("Triplicate", "deleteme"), "_")
AR <- separate(AR, "Triplicate", into = c("deleteme", "Triplicate"), -1)
AR <- AR[,-1]
AR[,c(4)] <- sapply(AR[,c(4)], as.numeric)
AR
#Remember to always remove S18


#### Site summary
Site_Sum <- Site_Summary[,c(1,2,3)]
colnames(Site_Sum)[1:3] <- c("Site", "Area", "Area_complex")

```

###FIGURES###
#Fig5
```{r N-s plot}

WALONG
WALONG$Element <- gsub("Acidity", "Acidity (mmol H+ /kg)", WALONG$Element) 

level_order <- c("SUL", "TZ", "TZ (inc)", "HYP")
Site_order <- c("S01", "", "", "", "S05", "", "", "", "S09", "", "", "", "S13", "", "", "", "S17", "", "", "", "S21", "")

ggplot(WALONG, aes(x=Site, y=mgKg, color = factor(Layer, level = level_order), group = factor(Layer, level = level_order)))+
  geom_point()+
  scale_x_discrete(labels = Site_order)+
  facet_wrap(.~ Element, nrow = 5, ncol = 3, scales = "free")+
  scale_color_manual("Layer",limits = c("SUL", "TZ", "TZ (inc)", "HYP"), labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c( "firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"))+
  theme(legend.position = c(0.5, 0.1),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.key.size = unit(2.5, "line"))

ggsave("Site_plots_water_sol.png", height = 40, width = 30, units = "cm")

```
#FIG1
#The water soluble boxplot for 3B FIG1
```{r BOXPLOTS all elements and acidity}

AS <- WALONG[WALONG$Element == "As",]
AS
median(AS[AS$Layer == "TZ (inc)",]$mgKg, na.rm = TRUE)
median(Water_sol_selected[Water_sol_selected$Layer == "2 Yes",]$As, na.rm = TRUE)

level_order <- c("SUL", "TZ", "TZ (inc)", "HYP")

WALONG
ggplot(WALONG, aes(x= factor(Layer, level = level_order), y=mgKg))+
  geom_boxplot( aes(color = Layer, group = Layer))+
  scale_y_log10(breaks = c(0.001, 0.01, 0.1, 1, 10, 100, 1000, 10000), labels = c(0.001,  0.01, 0.1, 1, 10, 100,  1000, 10000))+
  #geom_jitter(aes(color = Layer, group = Layer))+
  geom_signif(comparisons = list(c("SUL", "TZ"), c("SUL", "TZ (inc)"), c("SUL", "HYP"), c("TZ", "TZ (inc)"), c("TZ", "HYP"), c("TZ (inc)", "HYP")), map_signif_level =c(TRUE, " " = 0.001, " " =0.01, " " = 0.05), step_increase = 0.1, vjust = 0.7)+
  ylab("mg/Kg")+
  xlab("")+
  facet_wrap(.~Element, nrow = 7, ncol = 3, scales = "free")+
  theme(axis.title = element_text(size = 14),
        legend.text = element_text(size = 18),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 20),
        strip.text.x = element_text(size = 14),
        legend.position = c(0.5, 0.08),
        legend.key.size = unit(2.5, "line"))+
  scale_color_manual("Zone",limits = c("SUL", "TZ", "TZ (inc)", "HYP"), labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c( "firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"))

#Note, you have to manually input the significance stars in an image editor based on the results of the ANOVA further down (can't figure out how to get ggplot to do it for me)

ggsave("Water_Select_BOX_Log_X.png", width = 33, height = 45, unit = "cm")


```
#Fig6
#Acidity METALS 3B FIG6
```{r Acidity vs. vattenlösliga metaller}

#WALONG = for boxplots (Acidity is part of long). WALONG2 = for acidity vs. elements (Acidity is wide)

WALONG2

ggplot(na.omit(WALONG2), aes(x=Acidity, y=mgKg, color = Layer, shape = Layer))+
  geom_point()+
  facet_wrap(.~Element, nrow= 4, ncol = 3, scales = "free")+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Acidity (mmol H+/Kg)")+
  ylab("mg/Kg")+
  scale_color_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c("firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"), na.translate = F)+
    theme(axis.title = element_text(size = 14),
        legend.text = element_text(size = 13),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 15),
        strip.text.x = element_text(size = 14),
        legend.key.size = unit(1.8, "line"))+
  scale_shape_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c(16, 17, 18, 15))

ggsave("Acidity_Water.png", width = 30, height = 35, units = "cm")



```
#Fig2
#Total Sulfur vs. Acidity point correlation plot to 3B
```{r viz Acidity vs. AR S }

#Use AR S means for the transition layers
AR[AR$Layer=="2",]
Aqua_S_TZ_match <- AR[AR$Layer=="2",c(2,4)] %>%
  group_by(Site) %>%
  summarise_each(funs(mean(., na.rm = TRUE)))
#Remove S18
Aqua_S_TZ_match <- Aqua_S_TZ_match[-18,]

#merge acidity and aqua regia S
AR
Acidity
Acidity2 <- Acidity
Acidity2$Layer <- gsub("1 No", "1", Acidity2$Layer)
Acidity2$Layer <- gsub("2 No", "2", Acidity2$Layer)
Acidity2$Layer <- gsub("2 Yes", "2", Acidity2$Layer)
Acidity2$Layer <- gsub("3 Yes", "3", Acidity2$Layer)

Acidity2

Acidity_AR_S <- left_join(Acidity2, AR, by = c("Site", "Triplicate", "Layer"))
Acidity_AR_S <- left_join(Acidity_AR_S, Aqua_S_TZ_match, by = "Site")
Acidity_AR_S <- Acidity_AR_S %>%
  mutate(S.x = coalesce(S.x, S.y))

#Fix the Layer column
Acidity_AR_S
Acidity_AR_S <- Acidity_AR_S[,-c(5,8)]
Acidity_AR_S$Layer <- paste(Acidity_AR_S$Layer, Acidity_AR_S$Treatment)
Acidity_AR_S

#Remove the duplicates
Acidity_AR_S <- na.omit(Acidity_AR_S[!duplicated(Acidity_AR_S[,c("Site", "Layer", "Triplicate")]),])
#The S.y Is just the average for the TZ, to fill out the "empty" slots for TZ (After letjoin wihch didn't take it into account)
Acidity_AR_S
#Change S into percentage
Acidity_AR_S[,6] <- Acidity_AR_S[,6]/10000

#IDK. SOmething with the orders
facet_names <- list(
  '1 Non-inc' = 'OZ',
  '2 Non-inc' = 'Non-incubate TZ',
  '2 Incubated' = 'Incubated TZ',
  '3 Incubated' = 'RZ'
)
Acidity_AR_S$Treatment <- gsub("Non-inc", "Non-incubated", Acidity_AR_S$Treatment)
Acidity_AR_S
Acidity_AR_S$Treatment <- factor(Acidity_AR_S$Treatment, levels = c("Non-incubated","Incubated"))
Acidity_AR_S$Layer <- factor(Acidity_AR_S$Layer, levels = c("1 Non-inc", "2 Non-inc", "2 Incubated","3 Incubated"))

#Change the names
colnames(Acidity_AR_S)[6] <- "Sulfur"

ggplot(Acidity_AR_S, aes(x=Sulfur, y=Acidity, group = Layer, color = Layer))+
  geom_point()+
  facet_wrap(.~Treatment, ncol = 4, nrow = 1, scales = "free")+
  scale_color_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c( "firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"))+
  ylab("Acidity (mmol H+/Kg)")+
  xlab("Total Sulfur (%)")
ggsave("AR_SULFUR_VS_ACIDITY_Treatment_split_Acidity2.png", width = 20, height = 10, units = "cm")

Acidity_AR_S
cor.test(Acidity_AR_S[Acidity_AR_S$Layer == "1 Non-inc",]$Sulfur, Acidity_AR_S[Acidity_AR_S$Layer == "1 Non-inc",]$Acidity, method = "spearman")
cor.test(Acidity_AR_S[Acidity_AR_S$Layer == "2 Non-inc",]$Sulfur, Acidity_AR_S[Acidity_AR_S$Layer == "2 Non-inc",]$Acidity, method = "spearman")
cor.test(Acidity_AR_S[Acidity_AR_S$Layer == "2 Incubated",]$Sulfur, Acidity_AR_S[Acidity_AR_S$Layer == "2 Incubated",]$Acidity, method = "spearman")
cor.test(Acidity_AR_S[Acidity_AR_S$Layer == "3 Incubated",]$Sulfur, Acidity_AR_S[Acidity_AR_S$Layer == "3 Incubated",]$Acidity, method = "spearman")


```

#Fig 3
#S vs total S for Table 2
#Sidenote of S vs S ration #FIG 3. and #Table 2
```{r water S vs AR S}

Sulfur <- Water_sol_selected[,c(1:3,14,18)]
Sulfur <- left_join(Sulfur, Acidity_AR_S, by = c("Site","Layer", "Triplicate"))
#Sulfur is AR Sulfur, and S is water soluble S
Sulfur <- Sulfur[!duplicated(Sulfur[,c("Site", "Layer", "Triplicate")]),]
#Create the ratio between AR and water-sol S. Because AR S is in percentage, we need to divide by 10 000 to also put water sol in percentage
Sulfur$Ratio <- ((Sulfur$S/10000) / Sulfur$Sulfur)*100
Sulfur

#########
#THis is for Table 2

#Correlation test between water-sol sulfur (S) and Aqua regia sulfur (Sulfur)
#Does it matter if they are different units? One is mg/Kg one is percentage... It shouldn't matter, I think
cor.test(Sulfur[Sulfur$Layer == "3 Yes",]$S, Sulfur[Sulfur$Layer == "3 Yes",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "2 Yes",]$S, Sulfur[Sulfur$Layer == "2 Yes",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "2 No",]$S, Sulfur[Sulfur$Layer == "2 No",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "1 No",]$S, Sulfur[Sulfur$Layer == "1 No",]$Sulfur, method = "spearman")
#Conclusion, all are high (>0.5), with low p (<0.05), which means they correlate. Visible here:

ggplot(Sulfur, aes(x = S, y = Sulfur, color = Layer))+
  geom_point()

Sulfur
quantile(Sulfur[Sulfur$Layer == "3 Yes",]$Ratio, na.rm = TRUE)
quantile(Sulfur[Sulfur$Layer == "2 Yes",]$Ratio, na.rm = TRUE)
quantile(Sulfur[Sulfur$Layer == "2 No",]$Ratio, na.rm = TRUE)
quantile(Sulfur[Sulfur$Layer == "1 No",]$Ratio, na.rm = TRUE)


#This is the same as above, but instead of doing them one at a time, I've used pipe to make it into one calculation. Although it doesn't work for quantiles.
#S_S_Spearman <- S_S[,-c(1,3,7)] %>%
#  group_by(Layer, Treatment)%>%
#  summarise_each(funs(spearman_RHO = cor.test(S, Sulfur_AR, method = "spearman", exact = FALSE)$estimate, spearman_P = cor.test(S, Sulfur_AR, exact = FALSE, #method = "spearman")$p.val))


#####Key plots####

#Ratio vs. Aqua Regia sulfur
ggplot(Sulfur, aes(y=Ratio, x=Sulfur, color=Layer, shape = Treatment.x))+
  geom_point()+
  scale_x_log10()
ggsave("S_S_A.png", units = "cm", height=15, width = 15)

Sulfur
Sulfur[,c(3,5,7,8,9)]

Sulfur_long <- Sulfur[,c(3,5,7,8,9)]%>%
  gather(key = "Method", value = "value", -"Layer", -"Treatment.x", -"Ratio")
Sulfur_long

Sulfur_long$Method <- gsub("Sulfur", "Total Sulfur (mg/Kg)", Sulfur_long$Method) 
Sulfur_long$Method <- gsub("Acidity", "Acidity (mmol H+/Kg)", Sulfur_long$Method) 


##### THIS IS FIG 3 ############
Sulfur_long
ggplot(Sulfur_long, aes(x=value, y=Ratio, color=Layer))+
  geom_point()+
  facet_wrap(.~Method, ncol = 2, nrow = 1, scales = "free")+
  ylab("S(water) / S(total) * 100")+
  xlab("mmol H+/Kg                                                         mg/Kg ")+
  scale_color_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c("firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"), na.translate = F)+
  scale_shape_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c(16, 17, 18, 15))

ggsave("SRatios.png", units = "cm", height=10, width = 25)


cor.test(Sulfur[Sulfur$Layer == "3 Yes",]$Acidity, Sulfur[Sulfur$Layer == "3 Yes",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "2 Yes",]$Acidity, Sulfur[Sulfur$Layer == "2 Yes",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "2 No",]$Acidity, Sulfur[Sulfur$Layer == "2 No",]$Sulfur, method = "spearman")
cor.test(Sulfur[Sulfur$Layer == "1 No",]$Acidity, Sulfur[Sulfur$Layer == "1 No",]$Sulfur, method = "spearman")

```
#Fig4
#Fe vs S graph + spearman 3B FIG3
```{r R Fe vs S}

na.omit(Water_sol_selected[Water_sol_selected$Layer == "2 Yes",])
factor_level <- c("Non-incubated", "Incubated")

################# water sol vs water sol ##########

ggplot(Water_sol_selected, aes(x=S, y=Fe, color = Layer))+
  geom_point()+
  facet_wrap(.~factor(Treatment, levels = factor_level), nrow = 1, ncol = 3)+
   # scale_x_continuous(trans='log10', breaks = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000, 20000, 30000), labels = c(100, "", 300, "", "", "", "","","", 1000,"",3000,"", "","","","","", 10000,"", 30000))+
#  scale_y_continuous(trans='log10', breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000), labels = c(1, "", 3, "", "", "", "", "", "", 10, "", 30, "", "", "", "", "", "", 100, "", "300", "", "", "", "","","", 1000,"",3000,"", "","","","","", 10000))+
  scale_x_log10()+
  scale_y_log10()+
  scale_color_manual("Zone", labels = c("Oxidized zone", "Transition zone", "Transition zone (incubated)", "Reduced zone (incubated)"), values = c("firebrick", "darkgoldenrod2", "olivedrab3", "dodgerblue3"), na.translate = F)+
  xlab("Sulfur (mg/Kg)")+
  ylab("Iron (mg/Kg)")

ggsave("S_vsFe.png", width = 20, height = 10, units = "cm")

cor.test(Water_sol_selected[Water_sol_selected$Layer == "1 No",]$S, Water_sol_selected[Water_sol_selected$Layer == "1 No",]$Fe, method = "spearman")
cor.test(Water_sol_selected[Water_sol_selected$Layer == "2 No",]$S, Water_sol_selected[Water_sol_selected$Layer == "2 No",]$Fe, method = "spearman")
cor.test(Water_sol_selected[Water_sol_selected$Layer == "2 Yes",]$S, Water_sol_selected[Water_sol_selected$Layer == "2 Yes",]$Fe, method = "spearman")
cor.test(Water_sol_selected[Water_sol_selected$Layer == "3 Yes",]$S, Water_sol_selected[Water_sol_selected$Layer == "3 Yes",]$Fe, method = "spearman")



```

###Tables and statistics###
#THIS is the ANOVA tests for Water sol
```{r ANOVA}

#ANOVA

write.csv(Water_sol_selected, "Water_sol.csv")

#Log change the values first.

#TO get all the ANOVA p-values, just run this and then change teh element ("Mg" -> "S" -> "Ca", etc. for each desired element. I'm sure you can streamline it somehow, but I don't know how.)
TukeyHSD(aov(data = Water_sol_selected, log(Cu) ~ Layer))

Acidity
TukeyHSD(aov(data = Acidity, log(Acidity) ~ Layer))


```
#This is North vs South mann whitney U test
```{r water sol N vs S}

#MANN WHITNEY U TEST NORTH VS SOUTH
Site_Sum
colnames(Site_Sum)[2:4] <- c("Site", "Area", "Area_complex")
mann_W_test <- left_join(Water_sol_selected[! duplicated(Water_sol_selected),], Site_Sum, by = "Site")

mann_W_test_OZ <- lapply(mann_W_test[mann_W_test$Layer == "1 No",4:17], function(x){
  wilcox.test(x ~ mann_W_test[mann_W_test$Layer == "1 No",]$Area, exact = FALSE)$p.val
})
mann_W_test_TZ <- lapply(mann_W_test[mann_W_test$Layer == "2 No",4:17], function(x){
  wilcox.test(x ~ mann_W_test[mann_W_test$Layer == "2 No",]$Area, exact = FALSE)$p.val
})
mann_W_test_TZ_inc <- lapply(mann_W_test[mann_W_test$Layer == "2 Yes",4:17], function(x){
  wilcox.test(x ~ mann_W_test[mann_W_test$Layer == "2 Yes",]$Area, exact = FALSE)$p.val
})
mann_W_test_RZ <- lapply(mann_W_test[mann_W_test$Layer == "3 Yes",4:17], function(x){
  wilcox.test(x ~ mann_W_test[mann_W_test$Layer == "3 Yes",]$Area, exact = FALSE)$p.val
})

mann_W_test_OZ
mann_W_test_TZ
mann_W_test_TZ_inc
mann_W_test_RZ

write.csv(mann_W_test_OZ, "P_values_OZ.csv")
write.csv(mann_W_test_TZ, "P_values_TZ.csv")
write.csv(mann_W_test_TZ_inc, "P_values_TZ_inc.csv")
write.csv(mann_W_test_RZ, "P_values_RZ.csv")


```
#This is summary table for Water sol.
```{r water soluble TABLE}

Water_tab_sum <- Water_sol_selected[!duplicated(Water_sol_selected),-c(1,2,18)] %>%
  group_by(Layer)%>%
  summarise_each(funs(median(., na.rm = TRUE), min(., na.rm=TRUE), max(., na.rm = TRUE)))
Water_tab_sum

write.csv(Water_tab_sum, "Water_sol_sum.csv")

Acidity_tab_sum <- Acidity[! duplicated(Acidity[,1:4]), -c(1,3,4)] %>%
  group_by(Layer)%>%
  summarise_each(funs(median(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE)))
Acidity_tab_sum

```
#Water vs. HCL
```{r The percentage of water to HCL}

#THe percentage

HCL_upd
Water <- Water_sol_selected
Water$Layer <- gsub("1 No", "1", Water$Layer)
Water$Layer <- gsub("2 No", "2", Water$Layer)
Water$Layer <- gsub("2 Yes", "2", Water$Layer)
Water$Layer <- gsub("3 Yes", "3", Water$Layer)
Water
HCL_percentage <- within(merge(Water, HCL_upd, by = c("Site", "Triplicate", "Layer")), {
  Al <- Al.x/Al.y*100
  As <- As.x/As.y*100
  Cd <- Cd.x/Cd.y*100
  Co <- Co.x/Co.y*100
  Cr <- Cr.x/Cr.y*100
  Cu <- Cu.x/Cu.y*100
  Fe <- Fe.x/Fe.y*100
  Mn <- Mn.x/Mn.y*100
  Ni <- Ni.x/Ni.y*100
  Pb <- Pb.x/Pb.y*100
  S <- S.x/S.y*100
  Zn <- Zn.x/Zn.y*100
})[,c("Site", "Triplicate", "Layer", "Treatment", "Al", "As", "Cd", "Co", "Cr", "Cu", "Fe", "Mn", "Ni", "Pb", "S", "Zn")]
HCL_percentage

#Both water and HCL in the same long dataframe
WALONG
HCL_upd

HCLONG <- HCL_upd %>%
  gather(key = "Element", value = "mgKg", -"Triplicate", -"Site", -"Layer")

WALONG$Layer2 = WALONG$Layer
WALONG$Layer2 <- gsub("SUL", "1", WALONG$Layer2)
WALONG$Layer2 <- gsub("TZ \\(inc\\)", "2", WALONG$Layer2)
WALONG$Layer2 <- gsub("TZ", "2", WALONG$Layer2)
WALONG$Layer2 <- gsub("HYP", "3", WALONG$Layer2)

colnames(HCLONG)[3] <- "Layer2"
HCLONG <- left_join(WALONG, HCLONG, by = c("Site", "Triplicate", "Layer2", "Element"))
HCLONG
#x = Water, y = HCL

```
```{r water vs HCL - spearman HCL}

#ALL correlations
Spearman_HCL <- na.omit(HCLONG)%>%
  group_by(Layer, Element)%>%
  summarise(corr_koef = cor.test(mgKg.x, mgKg.y, na.action = "na.exclude", exact = FALSE, method = "spearman")$estimate, P_VAL = cor.test(mgKg.x, mgKg.y, na.action = "na.exclude", exact = FALSE, method = "spearman")$p.val)
Spearman_HCL

write.csv(Spearman_HCL, "3B_Water_HCL_SPEARMAN.csv")


```
```{r water vs HCL - quantiles}
HCLONG
HCL_percentage

HCL_percentage_long <- HCL_percentage %>%
  gather(key = "Element", value = "Percentage", -"Site", -"Triplicate", -"Layer", -"Treatment")
HCL_percentage_long

#THis is something
#quantile(Water_HCL_ZONES_avr[c(Water_HCL_ZONES_avr$Element == "Fe" & Water_HCL_ZONES_avr$Layer == "2 _ Yes"),]$Water_mgKg_mean / Water_HCL_ZONES_avr[c(Water_HCL_ZONES_avr$Element == "Fe" & Water_HCL_ZONES_avr$Layer == "2 _ Yes"),]$HCL_mgKg_mean * 100, na.rm = TRUE)

Water_Quantiles <- HCL_percentage_long[,-c(1,2)]%>%
  group_by(Layer, Treatment, Element)%>%
  summarise_each(funs("25q" = quantile(., 0.25, na.rm = TRUE), "50q" = quantile(., 0.50, na.rm = TRUE), "75q" = quantile(., 0.75, na.rm = TRUE)))
Water_Quantiles

write.csv(Water_Quantiles, "water_HCL_Q.csv")

# See https://stackoverflow.com/questions/11443053/output-of-quantile-to-a-data-frame for possible solution

```
```{r Water vs HCL}

HCL_upd

#COde changed from previously having HCL_all, to HCL_upd, since the previous didn't have half the detection values, the latter do.
HCL_OZ_RZ <- subset(na.omit(HCL_upd), Layer %in% c("1", "3"))
colnames(HCL_OZ_RZ)<- gsub("_HCL", "", colnames(HCL_OZ_RZ))
HCL_TZ <- subset(na.omit(HCL_upd), Layer %in% "2")
colnames(HCL_TZ)<- gsub("_HCL", "", colnames(HCL_TZ))

HCL_TZ  

#Make each long, so that you can merge an have two columns
Water_Long_merge <- Water_CLEAN_OZ_RZ %>%
  gather(key = Element, value = Water_mgKg, -"Site", -"Triplicate", -"Layer", -"Treatment")
HCL_Long_merge <- HCL_OZ_RZ %>%
  gather(key = Element, value = HCL_mgKg, -"Triplicate", -"Site", -"Layer")
Water_HCL_OZ_RZ <- left_join(Water_Long_merge, HCL_Long_merge, by =c("Triplicate", "Site", "Layer", "Element"))

#And the same with TZ
Water_CLEAN_TZ
HCL_TZ
TZ_Long_water <- na.omit(Water_CLEAN_TZ) %>%
  gather(key = Element, value = Water_mgKg, -"Site", -"Triplicate", -"Layer", -"Treatment")
#BUt I need to use average HCL first
HCL_TZ_average <- HCL_TZ[,-c(1,3)] %>%
  group_by(Site)%>%
  summarize_each(funs(mean= mean(., na.rm = TRUE)))
colnames(HCL_TZ_average) <- gsub("_mean", "", colnames(HCL_TZ_average))

HCL_TZ_average

TZ_Long_HCL <- HCL_TZ_average %>%
  gather(key = Element, value = HCL_mgKg, -"Site")


Water_HCL_TZ <- left_join(TZ_Long_water, TZ_Long_HCL, by = c("Site", "Element"))
Water_HCL_TZ

Water_HCL_OZ_RZ 
Water_HCL_TZ

Water_HCL_ZONES <- rbind(Water_HCL_OZ_RZ, Water_HCL_TZ)
Water_HCL_ZONES$Layer <- paste(Water_HCL_ZONES$Layer, "_", Water_HCL_ZONES$Treatment)
W_H_Z <- Water_HCL_ZONES[Water_HCL_ZONES$Element %in% c("Al", "As", "B", "Ba", "Be", "Ca", "Cd", "Co", "Cr", "Cu", "Fe", "K", "Mg", "Mn", "Mo", "Na", "Ni", "P", "Pb", "S", "Sr", "U", "V", "Zn"),]
W_H_Z 

W_H_Z[W_H_Z$Element == "Na",]


```
#Acidity
```{r Acidity AR S - statistical tests: spearman correlation, t-test and ANOVA}

#Subeset each soil zone into different dataframes to make the statistical tests easier
Acidity_Aqua_S_RZ <- subset(Acidity_Aqua_S[!duplicated(Acidity_Aqua_S[,c("Site", "Layer", "Triplicate", "Treatment")]),], Layer == "3 _ Incubated")
Acidity_Aqua_S_TZ_INC <- subset(Acidity_Aqua_S[!duplicated(Acidity_Aqua_S[,c("Site", "Layer", "Triplicate", "Treatment")]),], Layer == "2 _ Incubated")
Acidity_Aqua_S_TZ_NON <- subset(Acidity_Aqua_S[!duplicated(Acidity_Aqua_S[,c("Site", "Layer", "Triplicate", "Treatment")]),], Layer == "2 _ Non-inc")
Acidity_Aqua_S_OZ <- subset(Acidity_Aqua_S[!duplicated(Acidity_Aqua_S[,c("Site", "Layer", "Triplicate", "Treatment")]),], Layer == "1 _ Non-inc")

#Paired t-test to see if acidity and sulfur correlate overall
t.test(Acidity_Aqua_S$Acidity2, Acidity_Aqua_S$Sulfur, exact = FALSE, paired = TRUE)

#Spearman test between aque regia sulfur and acidity for each zone respectively
cor.test(Acidity_Aqua_S_RZ$Acidity2, Acidity_Aqua_S_RZ$Sulfur, method = "spearman", exact = FALSE)
cor.test(Acidity_Aqua_S_OZ$Acidity2, Acidity_Aqua_S_OZ$Sulfur, method = "spearman", exact = FALSE)
cor.test(Acidity_Aqua_S_TZ_INC$Acidity2, Acidity_Aqua_S_TZ_INC$Sulfur, method = "spearman", exact = FALSE)
cor.test(Acidity_Aqua_S_TZ_NON$Acidity2, Acidity_Aqua_S_TZ_NON$Sulfur, method = "spearman", exact = FALSE)

#ANOVA to see how they groups differ between soil zones
one_way <- aov(data = Acidity[!duplicated(Acidity[,c("Site", "Layer", "Triplicate", "Treatment")]),], log(Acidity) ~ Layer)
summary(one_way)
TukeyHSD(one_way)

```
```{r Acidity North to south Mann whitney U test}

Acidity_U <- left_join(Acidity, Site_Sum, by = "Site")
Acidity_U <- Acidity[,-c(3,4,7,8)] %>%
  group_by(Site, Layer)%>%
  summarize_each(funs(mean = mean(., na.rm = TRUE)))

Acidity_U <- left_join(Acidity_U, Site_Sum, by = "Site")

Acidity_U_OZ <- subset(Acidity_U, Layer %in% "1 _ Non-inc")
Acidity_U_TZ <- subset(Acidity_U, Layer %in% "2 _ Non-inc")
Acidity_U_TZ2 <- subset(Acidity_U, Layer %in% "2 _ Incubated")
Acidity_U_RZ <- subset(Acidity_U, Layer %in% "3 _ Incubated")

Acidity_U_OZ$Area <- as.factor(Acidity_U_OZ$Area)
Acidity_U_TZ$Area <- as.factor(Acidity_U_TZ$Area)
Acidity_U_TZ2$Area <- as.factor(Acidity_U_TZ2$Area)
Acidity_U_RZ$Area <- as.factor(Acidity_U_RZ$Area)

lapply(Acidity_U_OZ[4], function(x){
  wilcox.test(x ~ Acidity_U_OZ$Area, exact = FALSE)$p.val
})
lapply(Acidity_U_TZ[4], function(x){
  wilcox.test(x ~ Acidity_U_TZ$Area, exact = FALSE)$p.val
})
lapply(Acidity_U_TZ2[4], function(x){
  wilcox.test(x ~ Acidity_U_TZ2$Area, exact = FALSE)$p.val
})
lapply(Acidity_U_RZ[4], function(x){
  wilcox.test(x ~ Acidity_U_RZ$Area, exact = FALSE)$p.val
})


```

###Other###
```{r Appendix - AR vs Water sol}
Aqua
Aqua_select <- Aqua[-c(1,2,201:223),-c(1,2,4,7,8,51:57)]
colnames(Aqua_select)[1:4] <- c("Temp", "Site", "Area", "Layer")
Aqua_select <- separate(Aqua_select, "Temp", into = c("Triplicate", "deletemet"), "_")
Aqua_select <- separate(Aqua_select, "Triplicate", into = c("deleteme", "Triplicate"), -1)
Aqua_select <- Aqua_select[,-c(1,3)]
Aqua_select[,5:45] <- sapply(Aqua_select[,5:45], as.numeric)
Aqua_select <- Aqua_select[,c("Site", "Triplicate", "Area", "Layer", "Al", "As", "Cd", "Co", "Cr", "Cu", "Fe", "Mn", "Ni", "Pb", "S", "Zn")]
Aqua_select

Water_AR_compare <- Water_sol_selected
Water_AR_compare$Layer <- gsub("1 No", "1", Water_AR_compare$Layer)
Water_AR_compare$Layer <- gsub("2 No", "2", Water_AR_compare$Layer)
Water_AR_compare$Layer <- gsub("2 Yes", "2", Water_AR_compare$Layer)
Water_AR_compare$Layer <- gsub("3 Yes", "3", Water_AR_compare$Layer)

 
AppendixA <- within(merge(Water_AR_compare, Aqua_select, by=c("Site", "Triplicate", "Layer")), {
  Al <- Al.x/Al.y*100
  As <- As.x/As.y*100
  Cd <- Cd.x/Cd.y*100
  Co <- Co.x/Co.y*100
  Cr <- Cr.x/Cr.y*100
  Cu <- Cu.x/Cu.y*100
  Fe <- Fe.x/Fe.y*100
  Mn <- Mn.x/Mn.y*100
  Ni <- Ni.x/Ni.y*100
  Pb <- Pb.x/Pb.y*100
  S <- S.x/S.y*100
  Zn <- Zn.x/Zn.y*100
})[,c("Site", "Triplicate", "Layer", "Treatment", "Al", "As", "Cd", "Co", "Cr", "Cu", "Fe", "Mn", "Ni", "Pb", "S", "Zn")]
write.csv(AppendixA, "Water_sol_vs_AR.csv")

AppendixA <- AppendixA[,-c(1:2)] %>%
  group_by(Layer, Treatment)%>%
  summarize_each(funs("25Q" = quantile(., 0.25, na.rm = TRUE), "50Q" = quantile(., 0.50, na.rm = TRUE), "75Q" = quantile(., 0.75, na.rm = TRUE)))
write.csv(AppendixA, "Water_AR_quantiles.csv")

```
